---
title: "LAD calibration"
output: 
  html_document:
    toc: true
date: "2025-09-24"
author: "Nathéo BEAUCHAMP"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```

```{r}
library(targets)

library(dplyr)
library(tidyr)
library(data.table)

library(ggplot2)
library(cowplot)
library(patchwork)
library(ggforce)
library(legendry)
library(grid)
library(gridExtra)
library(ggnewscale)
library(scales)

library(BayesianTools)
library(coda)
library(nlme)
library(lme4)
library(merTools)
library(ggrepel)

library(sf)
```


```{r}
load("output/calib/outmods_20260605_localCompet_newSensorsFilter_continuousShadetol.RData")
load("output/comparison/outcomp_20260605_localCompet_newSensorsFilter_continuousShadetol.RData")
load("output/analysis/outanalysis_20260605_localCompet_newSensorsFilter_continuousShadetol.Rdata")
```

```{r}
tar_load(data_stands)
tar_load(init_db)
```

```{r}
scale_shadetol <- function(x) {
  
  mean_shadetol <- output_models_list$setup$`6`$shadetol_mean
  sd_shadetol <- output_models_list$setup$`6`$shadetol_sd
  
  (x - mean_shadetol) / sd_shadetol
}

scale_dbh <- function(x) {
  
  mean_dbh <- output_models_list$setup$`6`$dbh_mean
  sd_dbh <- output_models_list$setup$`6`$dbh_sd
  
  (x - mean_dbh) / sd_dbh
}

scale_compet_batot <- function(x) {
  
  mean_compet <- output_models_list$setup$`4`$compet_mean
  sd_compet <- output_models_list$setup$`4`$compet_mean
  
  (x - mean_compet) / sd_compet
}

scale_compet_bal <- function(x) {
  
  mean_compet <- output_models_list$setup$`5`$compet_mean
  sd_compet <- output_models_list$setup$`5`$compet_mean
  
  (x - mean_compet) / sd_compet
}

scale_compet_rci <- function(x) {
  
  mean_compet <- output_models_list$setup$`6`$compet_mean
  sd_compet <- output_models_list$setup$`6`$compet_mean
  
  (x - mean_compet) / sd_compet
}
```


# Models definition

```{r}
print(output_models_list$exp_design)

exp_design_short <- data.frame(
  id_model = 1:6,
  str_model = c("control by leaf group", 
                "+ shade tolerance",
                "+ interaction dbh",
                "+ interaction compet BATOT",
                "+ interaction compet BAL",
                "+ interaction compet RCI")
)
```


# Models comparison

## MAE and RMSE

```{r}
data_eval <- output_comparison_list$evaluation$indicators %>% 
  dplyr::bind_rows() %>% 
  dplyr::left_join(exp_design_short, by = "id_model") %>% 
  dplyr::group_by(i_rep, iteration) %>% 
  dplyr::mutate(
    MAE_control  = MAE[id_model == 1],
    RMSE_control = RMSE[id_model == 1],

    MAE_change  = 100 * (MAE  - MAE_control) / MAE_control,
    RMSE_change = 100 * (RMSE - RMSE_control) / RMSE_control
  ) %>% 
  dplyr::ungroup() %>% 
  dplyr::group_by(id_model, str_model) %>% 
  dplyr::summarise(
    
    MAE_lower = quantile(MAE, 0.025),
    MAE_median = median(MAE),
    MAE_upper = quantile(MAE, 0.975),
    
    MAE_change_lower = quantile(MAE_change, 0.025),
    MAE_change_median = median(MAE_change),
    MAE_change_upper = quantile(MAE_change, 0.975),
    
    RMSE_lower = quantile(RMSE, 0.025),
    RMSE_median = median(RMSE),
    RMSE_upper = quantile(RMSE, 0.975),
    
    RMSE_change_lower = quantile(RMSE_change, 0.025),
    RMSE_change_median = median(RMSE_change),
    RMSE_change_upper = quantile(RMSE_change, 0.975)
  ) %>% 
  dplyr::arrange(id_model)

data_eval
```


```{r}
mae_density <- output_comparison_list$evaluation$indicators %>% 
  dplyr::bind_rows() %>% 
  dplyr::left_join(exp_design_short, by = "id_model") %>% 
  
  ggplot(aes(x = MAE, color = str_model)) +
  geom_density() +
  theme_bw() +
  theme(legend.position = "none") +
  labs(color = "model ID")

mae_boxplot <- output_comparison_list$evaluation$indicators %>% 
  dplyr::bind_rows() %>% 
  dplyr::left_join(exp_design_short, by = "id_model") %>% 
  
  ggplot(aes(y = MAE, x = as.factor(id_model), color = str_model)) +
  geom_boxplot() +
  theme_bw() +
  theme(legend.position = "none") +
  xlab("model ID")

rmse_density <- output_comparison_list$evaluation$indicators %>% 
  dplyr::bind_rows() %>% 
  dplyr::left_join(exp_design_short, by = "id_model") %>% 
  
  ggplot(aes(x = RMSE, color = str_model)) +
  geom_density() +
  theme_bw() +
  theme(legend.position = "none") +
  labs(color = "model ID")

rmse_boxplot <- output_comparison_list$evaluation$indicators %>% 
  dplyr::bind_rows() %>% 
  dplyr::left_join(exp_design_short, by = "id_model") %>% 
  
  ggplot(aes(y = RMSE, x = as.factor(id_model), color = str_model)) +
  geom_boxplot() +
  theme_bw() +
  theme(legend.position = "top", legend.title = element_blank()) +
  xlab("model ID")

rmse_legend <- cowplot::get_legend(rmse_boxplot)
rmse_boxplot <- rmse_boxplot + theme(legend.position = "none")

cowplot::plot_grid(
  rmse_legend,
  cowplot::plot_grid(mae_density, mae_boxplot, nrow = 1),
  cowplot::plot_grid(rmse_density, rmse_boxplot, nrow = 1),
  ncol = 1, rel_heights = c(1, 2, 2)
)
```

## LOOIC

```{r}
data_comp <- output_comparison_list$comparison %>% 
  purrr::map(~data.frame(
    id_model = .x$id_model,
    i_rep = .x$i_rep,
    looic = .x[["LOOCV"]]["looic", "Estimate"],
    looic_se = .x[["LOOCV"]]["looic", "SE"],
    waic = .x[["WAIC"]]["waic", "Estimate"],
    waic_se = .x[["WAIC"]]["waic", "SE"]
  )) %>% 
  dplyr::bind_rows() %>% 
  dplyr::left_join(exp_design_short, by = "id_model") %>% 
  dplyr::arrange(waic) %>% 
  dplyr::mutate(
    looic_lower = looic - 1.96 * looic_se,
    looic_upper = looic + 1.96 * looic_se,
    waic_lower = waic - 1.96 * waic_se,
    waic_upper = waic + 1.96 * waic_se,
    
    delta_looic = looic - looic[id_model == 1],
    delta_waic = waic - waic[id_model == 1]
  ) %>% 
  
  dplyr::arrange(id_model)

data_comp
```

```{r}
waic_boxplot <- ggplot(data_comp, aes(y = waic, x = id_model, color = str_model)) +
  geom_point() +
  geom_pointrange(aes(ymin = waic_lower, ymax = waic_upper)) +
  theme_bw() +
  theme(legend.position = "none") +
  xlab("model ID") +
  ylab("WAIC")

looic_boxplot <- ggplot(data_comp, aes(y = looic, x = id_model, color = str_model)) +
  geom_point() +
  geom_pointrange(aes(ymin = looic_lower, ymax = looic_upper)) +
  theme_bw() +
  theme(legend.position = "top", legend.title = element_blank()) +
  xlab("model ID") +
  ylab("LOOIC")

looic_legend <- cowplot::get_legend(looic_boxplot)
looic_boxplot <- looic_boxplot + theme(legend.position = "none")

cowplot::plot_grid(
  looic_legend,
  cowplot::plot_grid(waic_boxplot, looic_boxplot, nrow = 1),
  ncol = 1, rel_heights = c(1, 4)
)
```



# Models convergence



# Prediciton of LAD

## Model 1: by leaf group

```{r}
output_analysis_list$output_params$`2` %>% 
      dplyr::select(id_set, 
                    p_intercept.broad, p_intercept.needle
                    ) %>% 
      tidyr::pivot_longer(!id_set, names_to = c("param", "leaf_group"), names_sep = "\\.") %>% 
  
  ggplot(aes(x = value, color = leaf_group)) +
  facet_wrap(~param, scales = "free") +
  geom_density() +
  theme_bw()
```

```{r}
output_analysis_list$output_params$`1` %>% 
  dplyr::mutate(
    lad_broad = exp(p_intercept.broad),
    lad_needle = exp(p_intercept.needle)
  ) %>% 
  dplyr::select(id_set, lad_broad, lad_needle) %>% 
  tidyr::pivot_longer(!id_set, 
                      names_to = "leaf_group", 
                      values_to = "lad",
                      names_prefix = "lad_") %>% 
  
  ggplot(aes(y = lad, x = leaf_group)) +
  geom_boxplot() +
  theme_bw() +
  xlab("Type of leaves") +
  ylab("LAD (m2/m3)")
```

## Model 2: consider species shade tolerance

```{r}
output_analysis_list$output_params$`2` %>% 
      dplyr::select(id_set, 
                    p_intercept.broad, p_intercept.needle, 
                    p_shadetol.broad, p_shadetol.needle
                    ) %>% 
      tidyr::pivot_longer(!id_set, names_to = c("param", "leaf_group"), names_sep = "\\.") %>% 
  
  ggplot(aes(x = value, color = leaf_group)) +
  facet_wrap(~param, scales = "free") +
  geom_density() +
  geom_vline(xintercept = 0, type = "dotted") +
  theme_bw()
```




```{r}
expand.grid(
  id_set = unique(output_analysis_list$output_params$`2`$id_set),
  leaf_group = c("needle", "broad"),
  shadetol = seq(1, 5, 0.1)
) %>% 
  dplyr::mutate(shadetol_scaled = scale_shadetol(shadetol)) %>% 
  dplyr::left_join(
    
    output_analysis_list$output_params$`2` %>% 
      dplyr::select(id_set, p_intercept.broad, p_intercept.needle, 
                    p_shadetol.broad, p_shadetol.needle) %>% 
      tidyr::pivot_longer(!id_set, names_to = c(".value", "leaf_group"), names_sep = "\\."),
    
    by = c("id_set", "leaf_group")
    
  ) %>% 
  
  # Predict LAD
  dplyr::mutate(
    lad = exp(p_intercept + shadetol_scaled * p_shadetol)
  ) %>% 
  
  # Compute summary statistics
  dplyr::group_by(leaf_group, shadetol) %>% 
  dplyr::summarise(
    lad_lower = quantile(lad, 0.025),
    lad_median = median(lad),
    lad_upper = quantile(lad, 0.975)
  ) %>% 
  
  ggplot(aes(y = lad_median, x = shadetol, color = leaf_group)) +
  
  geom_line() +
  geom_ribbon(aes(ymin = lad_lower, ymax = lad_upper,
                  fill = leaf_group),
              color = NA,
              alpha = 0.2) +
  theme_bw() +
  xlab("Species shade tolerance") +
  ylab("LAD (m2/m3)") +
  labs(color = "Type of leaves", fill = "Type of leaves")
```

## Model 3: add interaction with tree DBH

```{r}
output_analysis_list$output_params$`3` %>% 
      dplyr::select(id_set, 
                    p_intercept.broad, p_intercept.needle, 
                    p_shadetol.broad, p_shadetol.needle,
                    p_dbh.broad, p_dbh.needle,
                    p_shadetolXdbh.broad, p_shadetolXdbh.needle
                    ) %>% 
      tidyr::pivot_longer(!id_set, names_to = c("param", "leaf_group"), names_sep = "\\.") %>% 
  
  ggplot(aes(x = value, color = leaf_group)) +
  facet_wrap(~param, scales = "free") +
  geom_density() +
  geom_vline(xintercept = 0, type = "dotted") +
  theme_bw()
```

```{r}
expand.grid(
  id_set = unique(output_analysis_list$output_params$`3`$id_set),
  leaf_group = c("needle", "broad"),
  shadetol_group = c("shade intolerant", "shade tolerant"),
  dbh_cm = 10:75
) %>% 
  dplyr::mutate(
    shadetol = ifelse(shadetol_group == "shade intolerant", 2, 4.5),
    shadetol_scaled = scale_shadetol(shadetol),
    dbh_scaled = scale_dbh(dbh_cm)
  ) %>% 
  dplyr::left_join(
    
    output_analysis_list$output_params$`3` %>% 
      dplyr::select(id_set, 
                    p_intercept.broad, p_intercept.needle, 
                    p_shadetol.broad, p_shadetol.needle,
                    p_dbh.broad, p_dbh.needle,
                    p_shadetolXdbh.broad, p_shadetolXdbh.needle
                    ) %>% 
      tidyr::pivot_longer(!id_set, names_to = c(".value", "leaf_group"), names_sep = "\\."),
    
    by = c("id_set", "leaf_group")
    
  ) %>% 
  
  # Predict LAD
  dplyr::mutate(
    lad = exp(p_intercept + 
                shadetol_scaled * p_shadetol + 
                dbh_scaled * p_dbh + 
                shadetol_scaled * dbh_scaled * p_shadetolXdbh)
  ) %>% 
  
  # Compute summary statistics
  dplyr::group_by(leaf_group, shadetol_group, dbh_cm) %>% 
  dplyr::summarise(
    lad_lower = quantile(lad, 0.025),
    lad_median = median(lad),
    lad_upper = quantile(lad, 0.975)
  ) %>% 
  
  ggplot(aes(y = lad_median, x = dbh_cm, color = leaf_group)) +
  facet_wrap(~shadetol_group) +
  geom_line() +
  geom_ribbon(aes(ymin = lad_lower,
                  ymax = lad_upper,
                  fill = leaf_group),
              color = NA,
              alpha = 0.2) +
  theme_bw() +
  xlab("Tree DBH (cm)") +
  ylab("LAD (m2/m3)") +
  labs(color = "Type of leaves", fill = "Type of leaves")
```

## Models 4/5/6: add interaction with local competition

### Model 4: total basal area BATOT

```{r}
output_analysis_list$output_params$`4` %>% 
      dplyr::select(id_set, 
                    p_intercept.broad, p_intercept.needle, 
                    p_shadetol.broad, p_shadetol.needle,
                    p_dbh.broad, p_dbh.needle,
                    p_compet.broad, p_compet.needle,
                    p_shadetolXdbh.broad, p_shadetolXdbh.needle,
                    p_shadetolXcompet.broad, p_shadetolXcompet.needle,
                    p_dbhXcompet.broad, p_dbhXcompet.needle,
                    p_shadetolXdbhXcompet.broad, p_shadetolXdbhXcompet.needle
      ) %>% 
      tidyr::pivot_longer(!id_set, names_to = c("param", "leaf_group"), names_sep = "\\.") %>% 
  
  ggplot(aes(x = value, color = leaf_group)) +
  facet_wrap(~param, scales = "free") +
  geom_density() +
  geom_vline(xintercept = 0, type = "dotted") +
  theme_bw()
```

```{r}
expand.grid(
  id_set = unique(output_analysis_list$output_params$`4`$id_set),
  leaf_group = c("needle", "broad"),
  shadetol_group = c("shade intolerant", "shade tolerant"),
  dbh_cm = 10:75,
  compet_intensity = c("open", "dense")
) %>% 
  dplyr::mutate(
    shadetol = ifelse(shadetol_group == "shade intolerant", 2, 4.5),
    shadetol_scaled = scale_shadetol(shadetol),
    dbh_scaled = scale_dbh(dbh_cm),
    batot = case_match(compet_intensity, "open" ~ 15, "dense" ~ 45),
    batot_scaled = scale_compet_batot(batot)
  ) %>% 
  dplyr::left_join(
    
    output_analysis_list$output_params$`4` %>% 
      dplyr::select(id_set, 
                    p_intercept.broad, p_intercept.needle, 
                    p_shadetol.broad, p_shadetol.needle,
                    p_dbh.broad, p_dbh.needle,
                    p_compet.broad, p_compet.needle,
                    p_shadetolXdbh.broad, p_shadetolXdbh.needle,
                    p_shadetolXcompet.broad, p_shadetolXcompet.needle,
                    p_dbhXcompet.broad, p_dbhXcompet.needle,
                    p_shadetolXdbhXcompet.broad, p_shadetolXdbhXcompet.needle
                    ) %>% 
      tidyr::pivot_longer(!id_set, names_to = c(".value", "leaf_group"), names_sep = "\\."),
    
    by = c("id_set", "leaf_group")
    
  ) %>% 
  
  # Predict LAD
  dplyr::mutate(
    lad = exp(p_intercept + 
                shadetol_scaled * p_shadetol + 
                dbh_scaled * p_dbh + 
                batot_scaled * p_compet +
                shadetol_scaled * dbh_scaled * p_shadetolXdbh +
                shadetol_scaled * batot_scaled * p_shadetolXcompet +
                dbh_scaled * batot_scaled * p_dbhXcompet +
                shadetol_scaled * dbh_scaled * batot_scaled * p_shadetolXdbhXcompet)
  ) %>% 
  
  # Compute summary statistics
  dplyr::group_by(leaf_group, shadetol_group, dbh_cm, compet_intensity) %>% 
  dplyr::summarise(
    lad_lower = quantile(lad, 0.025),
    lad_median = median(lad),
    lad_upper = quantile(lad, 0.975)
  ) %>% 
  
  ggplot(aes(y = lad_median, x = dbh_cm, color = compet_intensity)) +
  facet_wrap(leaf_group~shadetol_group, scales = "free_y") +
  geom_line() +
  geom_ribbon(aes(ymin = lad_lower,
                  ymax = lad_upper,
                  fill = compet_intensity),
              color = NA,
              alpha = 0.2) +
  theme_bw() +
  xlab("Tree DBH (cm)") +
  ylab("LAD (m2/m3)") +
  labs(color = "Competition intensity")
```

### Model 5: basal area of larger trees BAL

```{r}
output_analysis_list$output_params$`5` %>% 
      dplyr::select(id_set, 
                    p_intercept.broad, p_intercept.needle, 
                    p_shadetol.broad, p_shadetol.needle,
                    p_dbh.broad, p_dbh.needle,
                    p_compet.broad, p_compet.needle,
                    p_shadetolXdbh.broad, p_shadetolXdbh.needle,
                    p_shadetolXcompet.broad, p_shadetolXcompet.needle,
                    p_dbhXcompet.broad, p_dbhXcompet.needle,
                    p_shadetolXdbhXcompet.broad, p_shadetolXdbhXcompet.needle
      ) %>% 
      tidyr::pivot_longer(!id_set, names_to = c("param", "leaf_group"), names_sep = "\\.") %>% 
  
  ggplot(aes(x = value, color = leaf_group)) +
  facet_wrap(~param, scales = "free") +
  geom_density() +
  geom_vline(xintercept = 0, type = "dotted") +
  theme_bw()
```


```{r}
expand.grid(
  id_set = unique(output_analysis_list$output_params$`5`$id_set),
  leaf_group = c("needle", "broad"),
  shadetol_group = c("shade intolerant", "shade tolerant"),
  dbh_cm = 10:75,
  compet_status = c("dominant", "suppressed")
) %>% 
  dplyr::mutate(
    shadetol = ifelse(shadetol_group == "shade intolerant", 2, 4.5),
    shadetol_scaled = scale_shadetol(shadetol),
    dbh_scaled = scale_dbh(dbh_cm),
    bal = case_match(compet_status, "dominant" ~ 0, "suppressed" ~ 30),
    bal_scaled = scale_compet_bal(bal)
  ) %>% 
  dplyr::left_join(
    
    output_analysis_list$output_params$`5` %>% 
      dplyr::select(id_set, 
                    p_intercept.broad, p_intercept.needle, 
                    p_shadetol.broad, p_shadetol.needle,
                    p_dbh.broad, p_dbh.needle,
                    p_compet.broad, p_compet.needle,
                    p_shadetolXdbh.broad, p_shadetolXdbh.needle,
                    p_shadetolXcompet.broad, p_shadetolXcompet.needle,
                    p_dbhXcompet.broad, p_dbhXcompet.needle,
                    p_shadetolXdbhXcompet.broad, p_shadetolXdbhXcompet.needle
                    ) %>% 
      tidyr::pivot_longer(!id_set, names_to = c(".value", "leaf_group"), names_sep = "\\."),
    
    by = c("id_set", "leaf_group")
    
  ) %>% 
  
  # Predict LAD
  dplyr::mutate(
    lad = exp(p_intercept + 
                shadetol_scaled * p_shadetol + 
                dbh_scaled * p_dbh + 
                bal_scaled * p_compet +
                shadetol_scaled * dbh_scaled * p_shadetolXdbh +
                shadetol_scaled * bal_scaled * p_shadetolXcompet +
                dbh_scaled * bal_scaled * p_dbhXcompet +
                shadetol_scaled * dbh_scaled * bal_scaled * p_shadetolXdbhXcompet)
  ) %>% 
  
  # Compute summary statistics
  dplyr::group_by(leaf_group, shadetol_group, dbh_cm, compet_status) %>% 
  dplyr::summarise(
    lad_lower = quantile(lad, 0.025),
    lad_median = median(lad),
    lad_upper = quantile(lad, 0.975)
  ) %>% 
  
  ggplot(aes(y = lad_median, x = dbh_cm, color = compet_status)) +
  facet_wrap(leaf_group~shadetol_group, scales = "free_y") +
  geom_line() +
  geom_ribbon(aes(ymin = lad_lower,
                  ymax = lad_upper,
                  fill = compet_status),
              color = NA,
              alpha = 0.2) +
  theme_bw() +
  xlab("Tree DBH (cm)") +
  ylab("LAD (m2/m3)") +
  labs(color = "Competition status")
```


### Model 6: light-based index RCI

```{r}
output_analysis_list$output_params$`6` %>% 
      dplyr::select(id_set, 
                    p_intercept.broad, p_intercept.needle, 
                    p_shadetol.broad, p_shadetol.needle,
                    p_dbh.broad, p_dbh.needle,
                    p_compet.broad, p_compet.needle,
                    p_shadetolXdbh.broad, p_shadetolXdbh.needle,
                    p_shadetolXcompet.broad, p_shadetolXcompet.needle,
                    p_dbhXcompet.broad, p_dbhXcompet.needle,
                    p_shadetolXdbhXcompet.broad, p_shadetolXdbhXcompet.needle
      ) %>% 
      tidyr::pivot_longer(!id_set, names_to = c("param", "leaf_group"), names_sep = "\\.") %>% 
  
  ggplot(aes(x = value, color = leaf_group)) +
  facet_wrap(~param, scales = "free") +
  geom_density() +
  geom_vline(xintercept = 0, type = "dotted") +
  theme_bw()
```

```{r}
expand.grid(
  id_set = unique(output_analysis_list$output_params$`6`$id_set),
  leaf_group = c("needle", "broad"),
  shadetol_group = c("shade intolerant", "shade tolerant"),
  dbh_cm = 10:75,
  compet_status = c("dominant", "suppressed")
) %>% 
  dplyr::mutate(
    shadetol = ifelse(shadetol_group == "shade intolerant", 2, 4.5),
    shadetol_scaled = scale_shadetol(shadetol),
    dbh_scaled = scale_dbh(dbh_cm),
    rci = case_match(compet_status, "dominant" ~ 0.4, "suppressed" ~ 1),
    rci_scaled = scale_compet_rci(rci)
  ) %>% 
  dplyr::left_join(
    
    output_analysis_list$output_params$`6` %>% 
      dplyr::select(id_set, 
                    p_intercept.broad, p_intercept.needle, 
                    p_shadetol.broad, p_shadetol.needle,
                    p_dbh.broad, p_dbh.needle,
                    p_compet.broad, p_compet.needle,
                    p_shadetolXdbh.broad, p_shadetolXdbh.needle,
                    p_shadetolXcompet.broad, p_shadetolXcompet.needle,
                    p_dbhXcompet.broad, p_dbhXcompet.needle,
                    p_shadetolXdbhXcompet.broad, p_shadetolXdbhXcompet.needle
                    ) %>% 
      tidyr::pivot_longer(!id_set, names_to = c(".value", "leaf_group"), names_sep = "\\."),
    
    by = c("id_set", "leaf_group")
    
  ) %>% 
  
  # Predict LAD
  dplyr::mutate(
    lad = exp(p_intercept + 
                shadetol_scaled * p_shadetol + 
                dbh_scaled * p_dbh + 
                rci_scaled * p_compet +
                shadetol_scaled * dbh_scaled * p_shadetolXdbh +
                shadetol_scaled * rci_scaled * p_shadetolXcompet +
                dbh_scaled * rci_scaled * p_dbhXcompet +
                shadetol_scaled * dbh_scaled * rci_scaled * p_shadetolXdbhXcompet)
  ) %>% 
  
  # Compute summary statistics
  dplyr::group_by(leaf_group, shadetol_group, dbh_cm, compet_status) %>% 
  dplyr::summarise(
    lad_lower = quantile(lad, 0.025),
    lad_median = median(lad),
    lad_upper = quantile(lad, 0.975)
  ) %>% 
  
  ggplot(aes(y = lad_median, x = dbh_cm, color = compet_status)) +
  facet_wrap(leaf_group~shadetol_group, scales = "free_y") +
  geom_line() +
  geom_ribbon(aes(ymin = lad_lower,
                  ymax = lad_upper,
                  fill = compet_status),
              color = NA,
              alpha = 0.2) +
  theme_bw() +
  xlab("Tree DBH (cm)") +
  ylab("LAD (m2/m3)") +
  labs(color = "Competition status", fill = "Competition status")
```


## Prediction with the RCI model

```{r}
cols <- c(
  "Shade-intolerant" = "#E69F00",
  "Shade-tolerant"   = "#0072B2"
)

fills <- cols

data_pred <- expand.grid(
  id_set = unique(output_analysis_list$output_params$`6`$id_set),
  leaf_group = c("needle", "broad"),
  shadetol_group = c("shade intolerant", "shade tolerant"),
  dbh_cm = 10:75,
  compet_status = c("dominant", "suppressed")
) %>%
  mutate(
    shadetol = ifelse(shadetol_group == "shade intolerant", 2, 4.5),
    shadetol_scaled = scale_shadetol(shadetol),
    dbh_scaled = scale_dbh(dbh_cm),
    rci = case_match(
      compet_status,
      "dominant" ~ 0.4,
      "suppressed" ~ 1
    ),
    rci_scaled = scale_compet_rci(rci)
  ) %>%

  left_join(

    output_analysis_list$output_params$`6` %>%
      select(
        id_set,
        p_intercept.broad, p_intercept.needle,
        p_shadetol.broad, p_shadetol.needle,
        p_dbh.broad, p_dbh.needle,
        p_compet.broad, p_compet.needle,
        p_shadetolXdbh.broad, p_shadetolXdbh.needle,
        p_shadetolXcompet.broad, p_shadetolXcompet.needle,
        p_dbhXcompet.broad, p_dbhXcompet.needle,
        p_shadetolXdbhXcompet.broad, p_shadetolXdbhXcompet.needle
      ) %>%
      pivot_longer(
        !id_set,
        names_to = c(".value", "leaf_group"),
        names_sep = "\\."
      ),

    by = c("id_set", "leaf_group")

  ) %>%

  mutate(

    lad = exp(
      p_intercept +
        shadetol_scaled * p_shadetol +
        dbh_scaled * p_dbh +
        rci_scaled * p_compet +
        shadetol_scaled * dbh_scaled * p_shadetolXdbh +
        shadetol_scaled * rci_scaled * p_shadetolXcompet +
        dbh_scaled * rci_scaled * p_dbhXcompet +
        shadetol_scaled * dbh_scaled * rci_scaled * p_shadetolXdbhXcompet
    )

  ) %>%

  group_by(
    leaf_group,
    shadetol_group,
    dbh_cm,
    compet_status
  ) %>%

  summarise(

    lad_lower = quantile(lad, 0.025),
    lad_median = median(lad),
    lad_upper = quantile(lad, 0.975),

    .groups = "drop"

  ) %>%

  mutate(

    leaf_group = factor(
      leaf_group,
      levels = c("needle", "broad"),
      labels = c("Needle-leaved", "Broad-leaved")
    ),

    shadetol_group = factor(
      shadetol_group,
      levels = c("shade intolerant", "shade tolerant"),
      labels = c("Shade-intolerant", "Shade-tolerant")
    )

  )


ggplot(
  data_pred,
  aes(
    x = dbh_cm,
    y = lad_median,
    colour = shadetol_group,
    fill = shadetol_group,
    linetype = compet_status
  )
) +

  facet_wrap(
    ~ leaf_group,
    nrow = 1
  ) +

  geom_ribbon(
    aes(
      ymin = lad_lower,
      ymax = lad_upper
    ),
    alpha = 0.12,
    colour = NA
  ) +

  geom_line(
    linewidth = 1.0
  ) +

  scale_colour_manual(values = cols) +
  scale_fill_manual(values = fills) +

  scale_linetype_manual(
    values = c(
      "dominant" = "solid",
      "suppressed" = "22"
    )
  ) +

  labs(
    x = "Stem diameter (cm)",
    y = expression(LAD~(m^2~m^{-3})),
    colour = "Shade tolerance",
    fill = "Shade tolerance",
    linetype = "Competition status"
  ) +

  theme_classic(base_size = 11) +

  theme(
    strip.background = element_blank(),
    strip.text = element_text(
      face = "bold",
      size = 11
    ),
    panel.spacing = unit(1.2, "lines"),
    legend.position = "top",
    legend.box = "vertical",
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 10),
    axis.title = element_text(size = 11),
    axis.text = element_text(size = 10)

  )
```

## Effect of competition/dbh on LAD prediction

```{r}
cols <- c(
  "Shade-intolerant" = "#E69F00",
  "Shade-tolerant"   = "#0072B2"
)

data_compare <- expand.grid(
  id_set = unique(output_analysis_list$output_params$`6`$id_set),
  leaf_group = c("needle", "broad"),
  shadetol_group = c("shade intolerant", "shade tolerant"),
  dbh_cm = c(10, 40, 70),
  compet_status = c("dominant", "suppressed")
) %>%
  mutate(

    shadetol = ifelse(
      shadetol_group == "shade intolerant",
      2,
      4.5
    ),

    shadetol_scaled = scale_shadetol(shadetol),

    dbh_scaled = scale_dbh(dbh_cm),

    rci = ifelse(
      compet_status == "dominant",
      0.4,
      1
    ),

    rci_scaled = scale_compet_rci(rci)

  ) %>%

  left_join(

    output_analysis_list$output_params$`6` %>%
      select(
        id_set,
        p_intercept.broad, p_intercept.needle,
        p_shadetol.broad, p_shadetol.needle,
        p_dbh.broad, p_dbh.needle,
        p_compet.broad, p_compet.needle,
        p_shadetolXdbh.broad, p_shadetolXdbh.needle,
        p_shadetolXcompet.broad, p_shadetolXcompet.needle,
        p_dbhXcompet.broad, p_dbhXcompet.needle,
        p_shadetolXdbhXcompet.broad,
        p_shadetolXdbhXcompet.needle
      ) %>%
      pivot_longer(
        !id_set,
        names_to = c(".value", "leaf_group"),
        names_sep = "\\."
      ),

    by = c("id_set", "leaf_group")

  ) %>%

  mutate(

    lad = exp(

      p_intercept +

      shadetol_scaled * p_shadetol +
      dbh_scaled * p_dbh +
      rci_scaled * p_compet +

      shadetol_scaled *
        dbh_scaled *
        p_shadetolXdbh +

      shadetol_scaled *
        rci_scaled *
        p_shadetolXcompet +

      dbh_scaled *
        rci_scaled *
        p_dbhXcompet +

      shadetol_scaled *
        dbh_scaled *
        rci_scaled *
        p_shadetolXdbhXcompet

    )

  ) %>%

  group_by(
    leaf_group,
    shadetol_group,
    dbh_cm,
    compet_status
  ) %>%

  summarise(
    lad_lower = quantile(lad, 0.025),
    lad_median = median(lad),
    lad_upper = quantile(lad, 0.975),
    .groups = "drop"
  ) %>%

  mutate(
    leaf_group = factor(
      leaf_group,
      levels = c("needle", "broad"),
      labels = c(
        "Needle-leaved",
        "Broad-leaved"
      )
    ),

    shadetol_group = factor(
      shadetol_group,
      levels = c(
        "shade intolerant",
        "shade tolerant"
      ),
      labels = c(
        "Shade-intolerant",
        "Shade-tolerant"
      )
    ),

    dbh_class = factor(
      dbh_cm,
      levels = c(10, 40, 70),
      labels = c(
        "Small tree\n(10 cm)",
        "Medium tree\n(40cm)",
        "Large tree\n(70 cm)"
      )
    ),
    
    compet_status = factor(
      compet_status,
      levels = c("suppressed", "dominant"),
      labels = c(
        "Suppressed",
        "Dominant"
      )
    )
  )


p2 <- ggplot(
  data_compare,
  aes(
    x = dbh_class,
    y = lad_median,
    colour = shadetol_group,
    shape = compet_status
  )
) +

geom_pointrange(
  aes(
    ymin = lad_lower,
    ymax = lad_upper
  ),
  position = position_dodge(width = 0.35),
  linewidth = 0.7
) +


facet_wrap(~leaf_group, nrow = 1) +

scale_colour_manual(values = cols) +
scale_shape_manual(values = c(16, 1)) +

labs(
  x = NULL,
  y = expression(LAD~(m^2~m^{-3})),
  colour = "Shade tolerance",
  shape = "Competition"
) +


theme_classic(base_size=11)+
theme(
  legend.position="top",
  legend.box="vertical",
  strip.background = element_blank(),
  strip.text = element_text(face="bold")
)

p2
```


# Crown volume as a mediator for leaf area density

```{r}
data_volume <- data_stands %>% 
  purrr::map(~.x$trees) %>% 
  dplyr::bind_rows(.id = "site") %>% 
  dplyr::left_join(init_db$plots, by = c("site" = "name"))
```

## Prediction of crown volume

```{r}
m0 <- lmer(
  log(volume_m3) ~ log(dbh_cm)*log(shadetol)*leaf_group +
    (1 | species) + (1 | origin/site),
  data = data_volume
)

m_rci <- lmer(
  log(volume_m3) ~ log(dbh_cm)*log(rci)*log(shadetol)*leaf_group +
    (1 | species) + (1 | origin/site),
  data = data_volume
)

m_batot <- lmer(
  log(volume_m3) ~ log(dbh_cm)*log(batot_local_m2ha)*log(shadetol)*leaf_group +
    (1 | species) + (1 | origin/site),
  data = data_volume
)

m_bal <- lmer(
  log(volume_m3) ~ log(dbh_cm)*log(bal_local_m2ha+0.0001)*log(shadetol)*leaf_group +
    (1 | species) + (1 | origin/site),
  data = data_volume
)

model_comp <- BIC(m0, m_rci, m_batot, m_bal) %>%
  as.data.frame() %>%
  mutate(
    delta_BIC = BIC - min(BIC)
  ) %>%
  arrange(BIC)

model_comp
```

```{r}
mod_volume <- m_rci

cols <- c(
  "Shade-intolerant" = "#E69F00",
  "Shade-tolerant"   = "#0072B2"
)

pred_grid <- expand.grid(
  dbh_cm = 10:75,
  compet_status = c("dominant", "suppressed"),
  leaf_group = c("needle", "broad"),
  shadetol_group = c("shade intolerant", "shade tolerant")
) %>%
  mutate(
    rci = ifelse(compet_status == "dominant", 0.4, 1),
    shadetol = ifelse(shadetol_group == "shade intolerant", 2, 4.5),
  ) %>% 
  
  mutate(

    ## population-level predictions
    volume_pred = exp(
      predict(
        mod_volume,
        newdata = .,
        re.form = NA
      ) +
      sigma(mod_volume)^2 / 2
    ),

    leaf_group = factor(
      leaf_group,
      levels = c("needle", "broad"),
      labels = c("Needle-leaved", "Broad-leaved")
    ),

    shadetol_group = factor(
      shadetol_group,
      levels = c("shade intolerant", "shade tolerant"),
      labels = c("Shade-intolerant", "Shade-tolerant")
    )
  )



ggplot(
  pred_grid,
  aes(
    x = dbh_cm,
    y = volume_pred,
    colour = shadetol_group,
    linetype = compet_status
  )
) +

  geom_line(linewidth = 1.1) +

  facet_wrap(
    ~leaf_group,
    nrow = 1
  ) +

  scale_colour_manual(values = cols) +

  scale_linetype_manual(
    values = c(
      dominant = "solid",
      suppressed = "22"
    )
  ) +

  labs(
    x = "Stem diameter (cm)",
    y = expression("Crown volume (" * m^3 * ")"),
    colour = "Shade tolerance",
    linetype = "Competition status"
  ) +

  theme_classic(base_size = 11) +

  theme(
    strip.background = element_blank(),

    strip.text = element_text(
      face = "bold",
      size = 11
    ),

    panel.spacing = unit(1.2, "lines"),

    legend.position = "top",
    legend.box = "vertical",

    legend.title = element_text(size = 10),
    legend.text = element_text(size = 10),

    axis.title = element_text(size = 11),
    axis.text = element_text(size = 10)
  )
```

## LAD vs crown volume

```{r}
predict_lad_site <- function(trees_df, params_df, site_name = NULL) {

  if(!is.null(site_name)) {
    
    site_origin <- init_db$plots$origin[init_db$plots$name == site_name]
    
    site_effects <- params_df %>%
      tidyr::pivot_longer(
        starts_with("p_z_site."),
        names_to = "site",
        values_to = "z_site",
        names_prefix = "p_z_site."
      ) %>%
      dplyr::filter(site == site_name)
    
    origin_effects <- params_df %>%
      tidyr::pivot_longer(
        starts_with("p_z_origin."),
        names_to = "origin",
        values_to = "z_origin",
        names_prefix = "p_z_origin."
      ) %>%
      dplyr::filter(origin == site_origin)
    
  }

  sp_effects <- params_df %>%
    tidyr::pivot_longer(
      starts_with("p_z_sp."),
      names_to = "species",
      values_to = "z_sp",
      names_prefix = "p_z_sp."
    )

 
  fixed_params <- params_df %>%
    dplyr::select(
      id_set,
      contains("sigma"),
      contains("p_intercept"),
      contains("p_shadetol"),
      contains("p_dbh"),
      contains("p_compet"),
      contains("p_shadetolX"),
      contains("p_dbhX"),
      contains("p_competX")
    ) %>%
    tidyr::pivot_longer(
      !c(id_set, contains("sigma")),
      names_to = c(".value", "leaf_group"),
      names_sep = "\\."
    )

  out <- tidyr::crossing(trees_df, params_df) %>%
    dplyr::left_join(sp_effects,   by = c("id_set", "species")) %>%
    dplyr::left_join(fixed_params, by = c("id_set", "leaf_group"))
  
  if (!is.null(site_name)) {
    out <- out %>% 
      dplyr::mutate(
      origin = site_origin,
      site = site_name
    ) 
  }
  
  out <- out %>% 

    dplyr::mutate(

      shadetol_scaled = scale_shadetol(shadetol),
      dbh_scaled = scale_dbh(dbh_cm),
      rci_scaled = scale_compet_rci(rci),

      sp_effect =
        z_sp * exp(p_sigma_sp_log),

      site_effect = ifelse(is.null(site_name), 0, 
                           z_origin * exp(p_sigma_origin_log) +
                             z_site   * exp(p_sigma_site_log)),

      lad = exp(
        p_intercept +
          sp_effect + 
          site_effect +
          shadetol_scaled * p_shadetol +
          dbh_scaled * p_dbh +
          rci_scaled * p_compet +
          shadetol_scaled * dbh_scaled * p_shadetolXdbh +
          shadetol_scaled * rci_scaled * p_shadetolXcompet +
          dbh_scaled * rci_scaled * p_dbhXcompet +
          shadetol_scaled * dbh_scaled * rci_scaled * p_shadetolXdbhXcompet
      )
    )

  # collapse inside stand
  out %>%
    dplyr::group_by(id_tree) %>%
    dplyr::summarise(
      lad = median(lad),
      .groups = "drop"
    )
}
```


```{r}
data_lad_per_ind <- purrr::imap(data_stands,
                                ~predict_lad_site(.x$trees, 
                                                  output_analysis_list$output_params$`6`)) %>% 
  dplyr::bind_rows(.id = "site") %>%
  left_join(
    data_stands %>%
      purrr::map(~.x$trees) %>%
      bind_rows(.id = "site"),
    by = c("site", "id_tree")
  ) %>% 
  dplyr::left_join(init_db$plots, by = c("site" = "name"))
```


```{r}
plot_df <- data_lad_per_ind %>%
  filter(!species %in% c("other_gymno", "other_angio")) %>% 
  dplyr::mutate(species = gsub("_", "\n", species))

species_order <- plot_df %>%
  dplyr::select(species, shadetol) %>% 
  dplyr::distinct() %>% 
  group_by(species) %>%
  arrange(shadetol) %>% 
  pull(species)

plot_df <- plot_df %>%
  mutate(species = factor(species, levels = species_order))

# ---------------------------------------------------------
# 1. Shared colour scale
# ---------------------------------------------------------
rc_scale <- scale_colour_viridis_c(
  option = "C",
  direction = -1,
  guide = guide_colourbar(
    direction = "horizontal",
    title.position = "top",
    title.hjust = 0.5,
    barwidth = 12,
    barheight = 0.8
  )
)

# ---------------------------------------------------------
# 2. Needle-leaved plot
# ---------------------------------------------------------
p_needle <- plot_df %>% 
  filter(leaf_group == "needle") %>% 
  
  ggplot(aes(
    x = volume_m3,
    y = lad,
    colour = rci
  )) +

  geom_point(alpha = 0.18, size = 1.1) +
  facet_wrap(~ species, scales = "free") +

  rc_scale +

  labs(
    x = expression("Observed crown volume (m"^3*")"),
    y = expression("Predicted LAD (m"^2*" m"^{-3}*")"),
    colour = "Competition (RCI)",
    title = "Needle-leaved species"
  ) +

  theme_classic(base_size = 11) +

  theme(
    strip.text = element_text(size = 7, face = "bold"),
    strip.background = element_blank(),
    panel.spacing = unit(0.9, "lines"),
    plot.title = element_text(hjust = 0.5),
    axis.title = element_text(size = 11, hjust = 0.5),
    axis.text.y = element_text(size = 9),
    axis.text.x = element_text(angle = 45, hjust = 0.8, size = 6),
    legend.position = "top",
    legend.direction = "horizontal"
  )

# ---------------------------------------------------------
# 3. Broad-leaved plot
# ---------------------------------------------------------
p_broad <- plot_df %>% 
  filter(leaf_group == "broad") %>% 
  
  ggplot(aes(
    x = volume_m3,
    y = lad,
    colour = rci
  )) +

  geom_point(alpha = 0.18, size = 1.1) +
  facet_wrap(~ species, scales = "free") +

  rc_scale +

  labs(
    x = expression("Observed crown volume (m"^3*")"),
    y = NULL,
    colour = "Competition (RCI)",
    title = "Broad-leaved species"
  ) +

  theme_classic(base_size = 11) +

  theme(
    strip.text = element_text(size = 7, face = "bold"),
    strip.background = element_blank(),
    panel.spacing = unit(0.9, "lines"),
    legend.position = "none",
    plot.title = element_text(hjust = 0.5),
    axis.title = element_text(size = 11),
    axis.text.y = element_text(size = 9),
    axis.text.x = element_text(angle = 45, hjust = 0.8, size = 6)
  )

# ---------------------------------------------------------
# 4. Get legend
# ---------------------------------------------------------
p_legend <- cowplot::get_legend(p_needle)
p_needle <- p_needle + theme(legend.position = "none")


# ---------------------------------------------------------
# 5. Combine 
# ---------------------------------------------------------
cowplot::plot_grid(
  p_legend,
  cowplot::plot_grid(
    p_needle, p_broad, nrow = 1
  ),
  rel_heights = c(1, 4),
  ncol = 1
)
``` 


# Crown expansion VS leaf packing strategies

```{r}
#############################################################
# DATA PREPARATION
#############################################################

pred_ladXvolume <- expand.grid(

  species = unique(data_volume$species),
  dbh_cm = c(25, 60),
  rci = c(0.4, 1)

) %>%

  dplyr::filter(!species %in% c("other_angio", "other_gymno")) %>%

  dplyr::left_join(
    data_volume %>%
      dplyr::select(species, leaf_group, shadetol) %>%
      dplyr::distinct(),
    by = "species"
  ) %>%

  dplyr::mutate(
    id_tree = row_number(),
    compet_status = ifelse(rci == 0.4, "dominant", "suppressed")
  ) %>%

  dplyr::left_join(
    predict_lad_site(., output_analysis_list$output_params$`6`),
    by = "id_tree"
  ) %>%

  dplyr::mutate(
    volume_pred = exp(
      predict(
        mod_volume,
        newdata = .,
        re.form = ~(1 | species)
      ) +
      sigma(mod_volume)^2 / 2
    )
  )

#############################################################
# ARROWS (dominant → suppressed)
#############################################################

arrows <- pred_ladXvolume %>%
  dplyr::select(species, dbh_cm, rci, volume_pred, lad) %>%
  tidyr::pivot_wider(
    names_from = rci,
    values_from = c(volume_pred, lad)
  )

#############################################################
# PLOT
#############################################################

ggplot() +

  # -------------------------------------------------------
  # arrows
  # -------------------------------------------------------
  geom_segment(
    data = arrows,
    aes(
      x = volume_pred_0.4,
      y = lad_0.4,
      xend = volume_pred_1,
      yend = lad_1,
      colour = species
    ),
    linewidth = 1,
    alpha = 0.85,
    arrow = arrow(length = unit(0.12, "cm"), type = "closed")
  ) +

  # -------------------------------------------------------
  # points (shape = DBH, fill = competition, colour = species)
  # -------------------------------------------------------
  geom_point(
    data = pred_ladXvolume,
    aes(
      x = volume_pred,
      y = lad,
      colour = species,
      fill = compet_status,
      shape = factor(dbh_cm)
    ),
    size = 3,
    stroke = 1
  ) +

  ###########################################################
  # SHAPE = DBH (circle vs triangle)
  ###########################################################

  scale_shape_manual(values = c(
    `25` = 21,   # small tree → circle
    `60` = 24    # large tree → triangle
  )) +

  ###########################################################
  # FILL = competition (dominant vs suppressed)
  ###########################################################

  scale_fill_manual(values = c(
    dominant = "white",     # empty symbol
    suppressed = "black"    # filled symbol
  )) +

  ###########################################################
  # KEEP species colour for borders + arrows
  ###########################################################

  guides(
    fill = guide_legend(override.aes = list(shape = 21)),
    colour = guide_legend(override.aes = list(fill = NA))
  ) +

  ###########################################################
  # LABELS
  ###########################################################

  labs(
    x = expression("Predicted crown volume (" * m^3 * ")"),
    y = expression("Leaf area density (LAD, m^2 ~ m^{-3})"),
    colour = "Species",
    fill = "Competition",
    shape = "DBH (cm)"
  ) +

  ###########################################################
  # THEME
  ###########################################################

  theme_classic(base_size = 12) +

  theme(
    legend.position = "top",
    legend.direction = "horizontal",

    legend.title = element_text(size = 10, face = "bold"),
    legend.text = element_text(size = 9),

    axis.title = element_text(size = 11),
    axis.text = element_text(size = 10),

    panel.spacing = unit(1.2, "lines"),

    strip.background = element_blank(),

    legend.box = "vertical"
  )
```

# Consequences on tree light interception

```{r}
out_energy_trees <- rbindlist(lapply(output_analysis_list$output_tree_light, function(x) {
  cbind(
    site = x$site_name,
    id_mod = x$id_mod,
    x$out_trees
  )
}), use.names = TRUE, fill = TRUE) %>% 
  
  tidyr::pivot_wider(names_from = "id_mod", values_from = c("e", "epot")) %>% 
  
  dplyr::mutate(
    diff_e = e_2 - e_6,
    diff_epot = epot_2 - epot_6,
    ratio_e = e_2 / e_6,
    ratio_epot = epot_2 / epot_6,
    ratio_scaled_e = (e_2 - e_6) / e_2,
    ratio_scaled_epot = (epot_2 - epot_6) / epot_2,
    logratio_e = log(e_2 / e_6),
    logratio_epot = log(epot_2 / epot_6),
  ) %>% 
  dplyr::group_by(site, id_tree) %>%
  dplyr::summarise(
    across(
      c(diff_e, diff_epot, ratio_e, ratio_epot, ratio_scaled_e, ratio_scaled_epot, logratio_e, logratio_epot),
      list(
        mean   = ~mean(.x, na.rm = TRUE),
        median = ~median(.x, na.rm = TRUE),
        q025   = ~quantile(.x, 0.025, na.rm = TRUE),
        q975   = ~quantile(.x, 0.975, na.rm = TRUE)
      ),
      .names = "{.col}_{.fn}"
    )
  ) %>% 
  dplyr::left_join(data_volume, by = c("site", "id_tree"))
```


```{r}
# ---------------------------------------------------------
# Shared theme
# ---------------------------------------------------------
base_theme <- theme_classic(base_size = 11) +
  theme(
    strip.background = element_blank(),
    strip.text = element_text(face = "bold", size = 7),

    legend.position = "top",
    legend.direction = "horizontal",
    legend.title = element_text(face = "bold", size = 10),
    legend.text = element_text(size = 9),

    axis.title = element_text(size = 11),
    axis.text.x = element_text(size = 7, angle = 45, hjust = 0.7),
    axis.text.y = element_text(size = 8),

    panel.spacing = unit(1, "lines")
  )

# ---------------------------------------------------------
# Filter dataset once
# ---------------------------------------------------------
df_energy <- out_energy_trees %>%
  filter(!species %in% c("other_gymno", "other_angio")) %>% 
  dplyr::mutate(species = gsub("_", "\n", species))

# ---------------------------------------------------------
# Shared colour scale
# ---------------------------------------------------------
rci_scale <- scale_colour_viridis_c(
  option = "C",
  direction = -1,
  name = "Competition (RCI)",
  guide = guide_colourbar(
    title.position = "top",
    title.hjust = 0.5,
    barwidth = 12,
    barheight = 0.8
  )
)

# ---------------------------------------------------------
# LEFT PANEL: epot
# ---------------------------------------------------------
p_epot <- ggplot(df_energy, aes(
  x = dbh_cm,
  y = logratio_epot_median,
  colour = rci
)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(alpha = 0.7, size = 1.6) +
  facet_wrap(~species, scales = "free_y") +
  rci_scale +
  labs(
    x = "DBH (cm)",
    y = expression("Potential intercepted energy (standardised)"),
    title = "a) Potential energy (no neighbours)"
  ) +
  base_theme

# ---------------------------------------------------------
# RIGHT PANEL: e
# ---------------------------------------------------------
p_e <- ggplot(df_energy, aes(
  x = dbh_cm,
  y = logratio_e_median,
  colour = rci
)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(alpha = 0.7, size = 1.6) +
  facet_wrap(~species, scales = "free_y") +
  rci_scale +
  labs(
    x = "DBH (cm)",
    y = expression("Intercepted energy (standardised)"),
    title = "b) Realised energy (with competition)"
  ) +
  base_theme +
  theme(legend.position = "none")

# ---------------------------------------------------------
# COMBINE (legend on top, above both panels)
# ---------------------------------------------------------
final_plot <- (p_epot + p_e) +
  plot_layout(ncol = 2, guides = "collect") &
  theme(
    legend.position = "top",
    legend.box = "horizontal"
  )

final_plot
```




```{r}
out_energy_trees %>% 
  dplyr::filter(!species %in% c("other_gymno", "other_angio")) %>% 
  
  ggplot(aes(y = logratio_epot_median, x = dbh_cm, color = rci)) +
  facet_wrap(~species, scales = "free_y") +
  
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  
  theme_classic() +
  theme(legend.position = "right") +
  scale_color_viridis_c(
    option = "C",
    direction = -1,
    guide = guide_colourbar(
      title.position = "top",
      title.hjust = 0.5
    )
  )
```

```{r}
out_energy_trees %>% 
  dplyr::filter(!species %in% c("other_gymno", "other_angio")) %>% 
  
  ggplot(aes(y = logratio_e_median, x = dbh_cm, color = rci)) +
  facet_wrap(~species, scales = "free_y") +
  
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  
  theme_classic() +
  theme(legend.position = "right") +
  scale_color_viridis_c(
    option = "C",
    direction = -1,
    guide = guide_colourbar(
      title.position = "top",
      title.hjust = 0.5
    )
  )
```


# Stand-level consequences on light on the ground

```{r}
cells_core <- purrr::imap(
  data_stands,
  function(stand, stand_name){

    #----------------------------------------
    # Convert cell centres to 5x5 m polygons
    #----------------------------------------
    cell_poly <- stand$cells %>%
      rowwise() %>%
      mutate(
        geometry = list(
          st_polygon(list(matrix(
            c(
              x_center - 2.5, y_center - 2.5,
              x_center + 2.5, y_center - 2.5,
              x_center + 2.5, y_center + 2.5,
              x_center - 2.5, y_center + 2.5,
              x_center - 2.5, y_center - 2.5
            ),
            ncol = 2,
            byrow = TRUE
          )))
        )
      ) %>%
      st_as_sf()

    #----------------------------------------
    # Keep cells intersecting the core polygon
    #----------------------------------------
    cell_poly$inside <- lengths(
      st_intersects(
        cell_poly,
        stand$core_polygon$sf
      )
    ) > 0

    cell_poly %>%
      filter(inside) %>%
      st_drop_geometry() %>%
      mutate(site = stand_name) %>%
      select(site, id_cell)

  }
) %>%
  bind_rows()
```

```{r}
out_energy_cells <- rbindlist(lapply(output_analysis_list$output_tree_light, function(x) {
  cbind(
    site = x$site_name,
    id_mod = x$id_mod,
    x$out_cells %>% dplyr::select(-e)
  )
}), use.names = TRUE, fill = TRUE) %>% 
  
  dplyr::right_join(cells_core, by = c("site", "id_cell")) %>%
  dplyr::select(site, id_set, id_cell, id_mod, pacl) %>%
  tidyr::pivot_wider(
    names_from = id_mod,
    values_from = pacl,
    names_prefix = "Model_"
  )
```

```{r}
ggplot(out_energy_cells, aes(Model_2, Model_6)) +
  
  geom_smooth(aes(colour = site), method = "lm", formula = y~x) +
  geom_abline(slope = 1, intercept = 0,
              linetype = "dashed", linewidth = 1) +
  
  coord_equal() +
  
  labs(
    x = "PACL with species-specific LAD model",
    y = "PACL with full LAD model"
  ) +
  
  theme_classic(base_size = 11) +
  theme(legend.position = "none")
```

